### 2022-04-06 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_betacov_mapping@23e2605f8bc7387ccc41cb2ad8ab052ac79e7759 🚀

>>>>

\hypertarget{host-distribution}{%
\subsection{Host distribution}\label{host-distribution}}

Chiroptera are an hyperdiverse group, distributed in a large part if the
world, and are an important reservoir for different strains of
betacoronaviruses (Drexler et al., 2014); this has attracted attention
to areas where high diversity of bats can be an important issue for
human health (Calisher et al., 2006). Accordingly, we collected the IUCN
rangemaps for known hosts of betacoronaviruses, to illustrate where
hotspots of host diversity are. These results are presented in Fig xx.a.
As per our current knowledge of which bats are hosts of
betacoronaviruses, these hotspots are primarily South-East Asia, parts
of Europe, and to a lesser extent sub-saharan Africa. Even the subset of
chiroptera that are hosts of betacoronaviruses fits the evolutionary
timeline of the group. Chiropterans can be classified as Microchiroptera
and macrochiroptera, where macrochiroptera have an older history from an
evolutionary perspective compared to macrochiroptera (Springer, 2013;
Teeling et al., 2005). South-East Asia has a high diversity of bats
(Kingston, 2010), and our results show that part of that diversity
includes betacoronavirus hosts. High density of hosts sharing the same
virus (albeit possibly different strains) calls into question the
evolution of the bat antiviral immune system and its co-evolution with
viruses, which may result in distinct immunological responses in
different area, as evidenced in other bat species (Banerjee et al.,
2020)

\hypertarget{viral-evolutionary-distinctiveness}{%
\subsection{Viral evolutionary
distinctiveness}\label{viral-evolutionary-distinctiveness}}

Higher host diversity may not result in a higher viral diversity; for
this reason, we quantified and mapped the evolutionary distinctiveness
of betacoronaviruses, based on \ldots. Viral evolutionary
distinctiveness largely tracks host diversity, particularly in southern
China but oddly not throughout the rest of southeast Asia, perhaps
indicating that many distinctive viruses remain to be discovered in this
region (an idea that is unsurprising given the growing realization,
around the emergence of SARS-CoV-2, that a unique lineage of similar
viruses are widespread in bats but still mostly undescribed). The most
distinct betacoronaviruses are found in South America, a region with a
comparatively lower number of hosts; this suggests that the South
American bat-betacoronvirus complex has been more isolated, and is
probably undergoing a different co-evolutionary dynamic. Alternatively,
this distinctiveness hostpot may be a product of under-sampling:
South-America is one of the places where the fewest betacoronaviruses
have been discovered (Anthony et al., 2017), and adding more viruses
would bring the distinctiveness of known sequences down. Previous work
has suggested the Americas may be a hotspot of both undiscovered bat
viruses in general (Olival) and coronavirus specifically (Anthony),
though not necessarily betacoronaviruses, and particularly not those in
clades with notable zoonotic potential (c.f. Anthony).

\hypertarget{geographic-mosaic-of-bat-betacoronavirus-risk}{%
\subsection{Geographic Mosaic of bat-betacoronavirus
risk}\label{geographic-mosaic-of-bat-betacoronavirus-risk}}

In order to turn the hypotheses based on the Geographic Mosaic Theory of
Coevolution into a measure of risk, we overlapped three components:
viral sharing, i.e.~the chance that two bats will share viruses overall;
Local Contribution to Beta Diversity, i.e.~the fact that a bat community
is compositionally unique compared to the average compositional
similarity across the entire system; finally, the phylogenetic
diversity, i.e.~how dispersed the bats in a location are within the tree
of life. These results are presented using an additive color mapping in
Figure xx, and lead to the definition of broad biogeographic regions of
risk, where the same color represents the same type of risk. Pairwise
maps of the three components are present in supplementary materials.

From the perspective of spillover risk, the most important combination
of factors is a high phylogenetic diversity of hosts with low viral
sharing; this, essentially, means that very different betacoronavirus
could co-exist within the same place. This is particularly the case
given that betacoronaviruses often evolve and even achieve host shifts
through recombination, which requires the co-occurrence of sufficiently
distinct viruses to be a major driver of emergence. In Fig. xx, this
corresponds to yellow to pale green areas, which are essentially limited
to South-Eastern Asia, and to some part of Sub-Saharan Africa. Adopting
a geographic mosaic theory perspective on risk, other regions of the
world are of lesser concern.

Available data on bat betacoronavirus spillover into humans (TP overlay
on the figure) is limited and circumstantial at best for these purposes,
but our risk maps suggest that the areas predicted by prior expectations
about host biogeography correspond loosely to those where previous
emergence events have been recorded. Areas with high bat diversity and
high turnover may facilitate the evolutionary radiation of viruses,
matching previous findings that the diversification of bat coronaviruses
is driven largely by host shifts (inter-genus or higher levels of
cross-species transmission) and, to a lesser degree, cospeciation and
sharing (intra-genus cross-species transmission; Anthony et al.~2017).
This diversification - while not an actual risk factor for spillover
itself - likely increases the random chance of a virus with the raw
genomic components required for the potential to infect humans.

\hypertarget{global-distribution-of-spillover-risk}{%
\subsection{Global distribution of spillover
risk}\label{global-distribution-of-spillover-risk}}

Based on the previous result, we extracted the yellow component of the
risk map (TP add methods), to provide a single measure of risk varying
between 0 and 1. This measure is presented in Fig. xxA. However, this
maps the potential risk, which must be weighed by the potential for
contacts with humans. As a proxy for this measure, we used the
proportion of build/urban land from the EarthEnv dataset: this is a
reasonable proxy for the density of humans per unit area, which
increases the probability of pathogen spread more widely (Hazarie et
al., 2021). Since human activity is required to amplify the frequency of
virus encounters and thus create areas of viral amplification, mapping
the potential risk against measures of land use is required to generate
a more actionable assessment of risk. This map is presented in Fig. xxB.
Most of South America and Europe are at low risk, as although densely
populated, settlements tend to be in areas with lower potential risk.
However, this mapping reveals that South-East Asia, the Indian
subcontinent, and parts of sub-Saharan Africa, are at high risk due to
the overlap between built areas and bat communities representing more
opportunities for cross-species transmission of betacoronaviruses.

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Driven by the need to understand the ecological factors involved in the
emergence of viral pathogens, we spatially mapped bat-betacoronavirus
interactions worldwide, using (i) a database of known betacov
hosts(Becker et al., 2020), and (ii) range maps for the hosts according
to IUCN (IUCN 2021). To reflect the fact that the risk posed by viruses
has many ecological origins, we quantified the phylogenetic diversity of
hosts, their compositional uniqueness, and the expected viral sharing.
Because these components of risk matter when contrasted to human
density, we compared them to a proxy, namely the proportion of each
pixel that is covered by urban or built land. This provides a synthetic
risk map, allowing to identifying of hotspots where the
bat-betacoronavirus system may originate viruses in humans. SE Asia is
one of the regions with the highest risk since, according to our
results, several of its conditions could increase the risk of
transmission of the virus.

Bats are found worldwide and are one of the most diverse groups among
mammals (Moratelli \& Calisher, 2015). Previous research (Anthony et
al., 2017; Mollentze \& Streicker, 2020) states that locally diverse bat
communities could maintain more viruses and hence, a higher probability
of having a pathogen that could represent a risk for human health. This
probability involves multiple factors, among which the relatedness of
hosts (which can make the jumps easier (Longdon et al., 2011; Mollentze
et al., 2020; Wolfe et al., 2007), and the overall tendency of hosts
within a locality to share viruses, which may limit viral diversity
because of within-host competition (Leeks et al., 2018; Sallinen et al.,
2020). Species richness, therefore, is not a sufficient measure of viral
risk. This is exemplified in our results, where both South America and
South-Eastern Asia have a high species richness of betacov hosts, but
only the latter region has a high risk. Specifically, because previous
studies propose that Asia is important when it comes to understanding
the evolutionary origin of various mammalian taxa (Beard C K, 1988).
Including bats (Yu et al., 2014), which could support the relationship
between evolutionary time and the development of an immune system with
characteristics that allow them to be better adapted to infection by
emerging viruses (Gorbunova et al., 2020; Irving et al., 2021) may be
related to a wide variety of diets (Jones et al., 2022; Moreno Santillán
et al., 2021; Banerjee et al., 2020; Schneeberger et al., 2013).

Our study focuses largely on the biogeography of hosts. Yet, we know
that viruses with high host plasticity, that is, the ability of a given
virus to adapt to various taxonomic orders and ecological groups
(Kreuder Johnson et al., 2015); are more likely to amplify viral
spillover, followed by secondary human-to-human transmission, and
geographical spread (Hazarie et al., 2021). High viral host plasticity
is an especially important trait for RNA viruses such as betacov
(Kreuder Johnson et al., 2015; Haddad et al., 2021). Indeed, our
analysis of viral sequences reveals that Latin America is a hotspot of
viral distinctiveness, suggesting that this part of the bats-betacov
system may be undergoing independent evolutionary dynamics (related
species sharing viruses that are different from the rest of the global
pool). The other hotspot of viral distinctiveness is S.E. Asia, in which
richness is high but sharing is low; this suggests a different type of
evolutionary dynamics (unrelated viruses coevolving with evolutionarily
distinct hosts, generating high diversity locally).

This diversity of hosts and how the exchange of viruses occurs between
species, is largely affected by the different environmental changes, as
the case of sarbecovirus bats reservoirs (Muylaert et al., 2021) where
they are affected by the area of the cave or the alteration of the
forest, which could result in modifications of host distribution.
Additionally, our results highlight the importance of Asia as a betacov
hotspot, which is consistent with recent studies (Muylaert et al.,
2021), where projections on this area suggest that new future events of
sarbecovirus viral exchange might be easily spread among species or
humans.

There are several factors that drive changes in the diversity of bats
(Alves et al., 2018), but human activities' effects on the ecosystem
(like modifications of land use) could significantly decrease it.
Therefore, it can be suggested that changes in the diversity of betacovs
in bats are linked to their biogeographic variation, and human
population density and other anthropogenic factors are decisive
moderators for its implications in public health. With the increase of
contact between humans and potential hosts, we also increase the risk of
emergence of novel diseases (Johnson et al., 2020), as previous studies
on RNA viruses suggest the importance of host phylogeography at the time
of virus dispersal (Gryseels et al., 2017).

One of these scenarios where interaction between bats and humans can
occur can be seed dispersal in tropical agroecosystems. It opens the
discussion of whether the fruits thrown by bats not only disperse seeds
but could also be a source of indirect interaction between viruses of
bat origin and humans (Deshpande et al., 2022) . This represents a
challenge for conservation strategies and disease ecology since we have
areas with potential zoonotic viruses and bat-human interaction.
However, it must still be taken into account the quantification of real
exposure from several scenarios, where there can be directly or
indirectly bat - human interaction.

Comparing scenarios of high viral exchange vs low viral exchange, open
the discussion to consider if the best scenario is where viruses easily
adapted to multiple hosts but with low virulence or easily ignored by
the immune system of the host, or where we have viruses specialized to a
specific host, but highly virulent when invade a new host. Accordingly,
the understanding of viral-host interactions from a taxonomic and
phylogenetic contributes to improving zoonoses surveillance programs.

\textbf{Acknowledgements}: We acknowledge that this study was conducted
on land within the traditional unceded territory of the Saint Lawrence
Iroquoian, Anishinabewaki, Mohawk, Huron-Wendat, and Omàmiwininiwak
nations. This work was supported by funding to the Viral Emergence
Research Initiative (VERENA) consortium including NSF BII 2021909 and a
grant from Institut de Valorisation des Données (IVADO). This research
was enabled in part by support provided by Calcul Québec
(www.calculquebec.ca) and Compute Canada (www.computecanada.ca). NF is
funded by the NSERC BIOS² CREATE program. TP and NF are funded by the
Courtois Foundation. RLM was supported by Bryce Carmine and Anne Carmine
(née Percival), through the Massey University Foundation.
